---
title: "Neuro analyses, ICD-9 code"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: false
    code_download: true
    highlight: tango
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(rcartocolor)
library(forcats)
library(purrr)
library(DT)
library(cowplot)

load('data/processed-data.Rdata')
theme_set(theme_bw() + theme(legend.title = element_blank()))
source('utils.R')
```

## Prevalence analysis

Calculate percentages:

```{r fig.height=9}
code_prevalence <- diag_icd_9 %>% 
  mutate(time = fct_relevel(time, c('After admission', 'Before admission'))) %>% 
  select(- contains('ICD10_')) %>% 
  group_by(siteid, time, icd, `Neurological Disease Category`, full_icd) %>% 
  mutate(percent_pats_site = num_patients_icd/num_patients_all,
         siteid = as.factor(siteid)) %>% 
  ungroup() %>% 
  group_by(Country, time, icd, `Neurological Disease Category`, full_icd) %>% 
  mutate(percent_pats_country = sum(num_patients_icd)/all_pats_country) %>% 
  ungroup()
```

```{r include=FALSE}
# diag_icd_9 %>% 
#   count(siteid, icd, `Neurological Disease Category`) %>% 
#   pull(icd) %>% 
#   unique() %>% 
#   length()
# 
# check = diag_icd_9 %>%
#   count(icd, `Neurological Disease Category`)
diag_icd_9 %>% 
  count(siteid, Country, num_patients_all)

diag_icd_9 %>% 
  count(siteid, Country, num_patients_all) %>% 
  filter(Country == 'Italy')

diag_icd_9 %>% filter(
  siteid %in% intersect(unique(diag_icd_9$siteid), unique(diag_icd_10$siteid))
) %>% 
  select(Country, siteid) %>% 
  distinct()

diag_icd_9 %>% 
  filter(Country == 'USA') %>% 
  group_by(siteid) %>% 
  summarise(site_pats_9 = sum(num_patients_icd), .groups = 'drop') %>% 
  left_join(diag_icd_10 %>% 
  filter(siteid %in% diag_icd_9$siteid) %>% 
  group_by(siteid) %>% 
  summarise(site_pats_10 = sum(num_patients_icd), .groups = 'drop'),
  by = 'siteid') %>% 
  pivot_longer(- siteid) %>% 
  ggplot(aes(x = siteid, y = value, fill = name)) +
  geom_col(position = 'dodge')
```


```{r fig.height=4, fig.width=10}
diff_code <- code_prevalence %>% 
  select(location = siteid, icd, `Neurological Disease Category`, time, percent_pats_site) %>% 
  pivot_wider(names_from = time, values_from = percent_pats_site, 
              id_cols = c(location, icd, `Neurological Disease Category`), 
              values_fill = list(percent_pats_site = 0)) %>% 
  mutate(percent_diff = `After admission` - `Before admission`,
         loc_type = 'Site', location = tolower(location))

diff_code_country <- code_prevalence %>% 
  select(location = Country, icd, `Neurological Disease Category`, time, percent_pats_country) %>% 
  distinct() %>% 
  pivot_wider(names_from = time, values_from = percent_pats_country, 
              id_cols = c(location, icd, `Neurological Disease Category`), 
              values_fill = list(percent_pats_country = 0)) %>% 
  mutate(percent_diff = `After admission` - `Before admission`,
         loc_type = 'Country')

diff_code_loc <- diff_code %>% 
  bind_rows(diff_code_country)

diff_heat <- diff_code_loc %>% 
  ggplot(aes(y = icd, x = location, fill = percent_diff)) +
  geom_tile() +
  labs(y = NULL, x = NULL, fill = 'After - Before') +
  scale_x_discrete(drop = F) + 
  scale_fill_gradient2(#009392,#39b185,#9ccb86,#e9e29c,#eeb479,#e88471,#cf597e
    low = '#798234',
    high = '#cf597e',
    labels = scales::percent_format(accuracy = 1)) +
  facet_grid(rows = vars(`Neurological Disease Category`), 
             cols = vars(fct_rev(loc_type)), space = 'free', scale = 'free') +
  heat_theme_bottom() +
  theme(panel.spacing.x = unit(10, "points"),
        plot.margin = unit(c(1,8,0.5,12), "lines"),
        axis.text.x = element_text(hjust = 1)) +
  # theme() +
  NULL
diff_heat

 # ggsave('figs/icd9_diff_heatmap.png', diff_heat, height = 4, width = 8)
```



### Hypothesis testing

```{r}
my_t <- function(icdi){
  diff_icd <- diff_code %>% filter(icd == icdi)
  data.frame(
    icd = icdi,
    t.test(diff_icd$percent_diff) %>% 
    broom::tidy()
  )
}
prevalence_stats <- unique(diff_code$icd) %>% 
  lapply(my_t) %>% 
  bind_rows()
prevalence_stats$p_value_holm <-  p.adjust(prevalence_stats$p.value, 'holm')
prevalence_stats$p_value_bh <- p.adjust(prevalence_stats$p.value, 'BH')
prevalence_stats %>% 
  filter(p_value_bh < 0.05)

```


### Prevalence change table

```{r}
prevalence_stats %>% 
  arrange(desc(estimate)) %>% 
  mutate(ci = paste0('(', round(conf.low*100, 2), '%, ', 
                     round(conf.high*100, 2), '%)')) %>% 
  select(- c(parameter, method, alternative, conf.low, conf.high)) %>% 
  datatable(rownames = FALSE, filter = 'top') %>% 
  formatRound(c('statistic'), 1) %>%
  formatPercentage('estimate', 1) %>% 
  formatSignif(c('p.value', 'p_value_holm', 'p_value_bh'), 3) %>%
  {.}

prevalence_stats %>% 
  write_csv('results/icd9_prevalence_stats.csv')
```


### Compute confidence inverval 
of the mean proportion of patients diagnosed with each ICD

```{r}
alpha_threshold <- qnorm(0.975)

ci_prevalence <- code_prevalence %>% 
  group_by(time, full_icd) %>%
  add_count() %>%
  summarise(
    mean_prop = mean(percent_pats_site, na.rm = T),
    sd_prob = sd(percent_pats_site, na.rm = T),
    n = mean(n),
    me_prop = alpha_threshold * sd_prob / sqrt(n)
  ) %>%
  ungroup()
```



```{r fig.width=10, fig.height=5}
icd9_time <- diag_icd_9 %>% 
  mutate(time = fct_relevel(time, c('After admission', 'Before admission'))) %>% 
  group_by(full_icd, time) %>% 
  summarise(pats_icd9_time = sum(num_patients_icd, na.rm = T), .groups = 'drop') 

sorted_icds <- code_prevalence %>% 
  ungroup() %>% 
  select(time, percent_pats_site, full_icd, siteid) %>% 
  pivot_wider(names_from = time, values_from = percent_pats_site, values_fill = list(percent_pats_site = 0)) %>% 
  group_by(full_icd) %>% 
  summarise(after = mean(`After admission`, na.rm = T),
         before = mean(`Before admission`, na.rm = T),
         diff_prev = after - before,
         .groups = 'drop') %>% 
  arrange(diff_prev) %>% 
  pull(full_icd)

total_icd <- icd9_time %>% 
  ggplot(aes(x = pats_icd9_time, y = fct_relevel(full_icd, sorted_icds), fill = time)) +
  scale_fill_carto_d(palette = 4, guide = guide_legend(reverse = TRUE)) +
  geom_col(position = 'dodge') +
  theme_minimal() +
  scale_x_reverse(expand = expansion(add = c(0,0))) +
  scale_y_discrete(labels = NULL) +
  theme(legend.position = c(0.3, 0.15),
        legend.title = element_blank(),
        legend.key.height = unit(4, 'mm'),
        plot.margin = margin(t = 1, r = 0.5, l = 0.5, unit = 'lines')) + 
  labs(x = 'Total number of patients at all sites', y = NULL)

percent_icd <- 
  ci_prevalence %>% 
  ggplot(aes(group = time)) +
  ggstance::geom_pointrangeh(
    aes(x = mean_prop,
        y = fct_relevel(full_icd, sorted_icds),
        xmin = mean_prop - me_prop,
        xmax = mean_prop + me_prop,
        color = time),
    position = position_dodge(width = 0.3), 
    stroke = 0.1, fatten = 3, size = 0.7  
  ) +
  scale_color_carto_d(palette = 4, guide = NULL) +
  theme_minimal() +  
  theme(legend.position = c(0.75, 0.1),
        legend.title = element_blank(),
        axis.text.y = element_text(hjust = 0.5),
        plot.margin = margin(t = 1, unit = 'lines')) + 
  scale_x_continuous(expand = expansion(add = c(0, 0.03)),
                     labels = scales::percent_format(accuracy = 1)) +
  labs(x = 'Proportion of patients each site', y = NULL)

icd9_prevalence_plots <- cowplot::plot_grid(total_icd, percent_icd, ncol = 2,
                                           rel_widths = c(1, 2.5))
icd9_prevalence_plots
# ggsave('figs/icd9_prevalence.png', icd9_prevalence_plots, height = 5, width = 10)
```

```{r}
plot_grid(diff_heat, icd9_prevalence_plots, rel_heights = c(1, 0.85),
          ncol = 1, labels = 'AUTO') %>%
  ggsave('figs/icd9_prevalence_AB.png', ., height = 8, width = 9.7)

plot_grid(diff_heat, icd9_prevalence_plots, rel_heights = c(1, 0.85),
          ncol = 1, labels = 'AUTO') %>%
  ggsave('figs/tiffs/efig_4.tiff', ., height = 8, width = 9.7, dpi = 300)
```

```{r fig.height=9}
before_code_prev <- diff_code_loc %>% 
  ggplot(aes(y = icd, x = location, fill = `Before admission`)) +
  geom_tile() +
  labs(y = NULL, x = NULL) +
  #f6d2a9,#f5b78e,#f19c7c,#ea8171,#dd686c,#ca5268,#b13f64
  scale_fill_gradient(
    low = '#d1eeea', high = '#2a5674',
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, max(diff_code_loc$`After admission`)),
    guide = guide_legend(override.aes = list(fill = "white"))) +
  scale_x_discrete(drop = F) + 
  facet_grid(rows = vars(`Neurological Disease Category`), 
             cols = vars(fct_rev(loc_type)), space = 'free', scale = 'free') +
  heat_theme_top() +
  NULL

after_code_prev <- diff_code_loc %>% 
  ggplot(aes(y = icd, x = location, fill = `After admission`)) +
  geom_tile() +
  labs(y = NULL, x = NULL, fill = 'Patient % per
site/country') +
  scale_x_discrete(drop = F) + 
  scale_fill_gradient(low = '#d1eeea', high = '#2a5674',
                      labels = scales::percent_format(accuracy = 1),
                      limits = c(0, max(diff_code_loc$`After admission`))) +
  facet_grid(rows = vars(`Neurological Disease Category`), 
             cols = vars(fct_rev(loc_type)), space = 'free', scale = 'free') +
  heat_theme_bottom() +
  theme(legend.title = element_text(hjust = 0)) +
  NULL

heats <- cowplot::plot_grid(before_code_prev, after_code_prev, ncol = 1,
                            rel_heights = c(1, 1.15),
                   labels = c('A. Before admission', 'B. After admission'))
heats
ggsave('figs/icd9_heatmap.png', heats, height = 10, width = 7)
ggsave('figs/tiffs/efig_3.tiff', heats, height = 10, width = 7, dpi = 300)
```


```{r}
diff_code %>% 
  filter(icd == 'R41') %>% 
  ungroup() %>% 
  count(percent_diff > 0)

diff_code %>% 
  filter(icd == 'G93') %>% 
  ungroup() %>% 
  count(percent_diff > 0)

diff_code %>% 
  filter(icd == 'R42') %>% 
  ungroup() %>% 
  count(percent_diff > 0)
```


# Severity descriptive statistics

What ICD code has the most number severe patients?
Sites with more severe patients?

```{r fig.height=9}
severe_code <- diag_icd_9 %>% 
  mutate(percent_severe_site = num_patients_ever_severe_icd1/num_patients_icd,
         siteid = as.factor(siteid))

severe_bef_heat <- severe_code %>% 
  filter(time == 'Before admission') %>% 
  ggplot(aes(y = icd, x = siteid, fill = percent_severe_site)) +
  geom_tile() +
  labs(y = NULL, x = NULL, fill = 'Severe % per ICD') +
  #3d5941,#778868,#b5b991,#f6edbd,#edbb8a,#de8a5a,#ca562c
  scale_fill_gradient(
    # low = '#f6d2a9', high = '#b13f64',
    low = 'white', high = '#ca562c',
    guide = guide_legend(override.aes = list(fill = "white")),
    ) +
  scale_x_discrete(drop = F) + 
  facet_grid(rows = vars(`Neurological Disease Category`), space = 'free', scale = 'free') +
  heat_theme_top() +
  NULL

severe_aft_heat <- severe_code %>% 
  filter(time == 'After admission') %>% 
  ggplot(aes(y = icd, x = siteid, fill = percent_severe_site)) +
  geom_tile() +
  labs(y = NULL, x = NULL, fill = 'Severe % per ICD') +
  scale_x_discrete(drop = F) + 
  #3d5941,#778868,#b5b991,#f6edbd,#edbb8a,#de8a5a,#ca562c
  scale_fill_gradient(
    low = 'white', high = '#ca562c',
    labels = scales::percent_format(accuracy = 1)) +
  facet_grid(rows = vars(`Neurological Disease Category`), space = 'free', scale = 'free') +
  heat_theme_bottom() +
  NULL

severe_heats <- cowplot::plot_grid(severe_bef_heat, severe_aft_heat, ncol = 1,
                            rel_heights = c(1, 1.15),
                   labels = c('A. Before admission', 'B. After admission'))
severe_heats
ggsave('figs/icd9_severe_heatmap.png', severe_heats, height = 10, width = 7)
```



# Severity enrichment analysis

## Null hypothesis 
For each ICD code, the proportion of severe patients who were diagnosed with that ICD code is similar to the proportion of never-severe patients who were diagnosed with that same ICD code.

For the sake of simplicity in this notebook, we're going to denote patients whose symptoms have been categorized as severe (based on respiratory status +/- requiring ICU) at least once as *severe patients*, and patients whose symptoms have NEVER been categorized as severe as *non-severe patients*.

For each ICD code, we computed the expected number of *severe patients* by multiplying the proportion of *non-severe patients* who were diagnosed with that code with the total number of *severe patients*.
We performed an enrichment analysis to examine the difference of *severe patients* proportions across ICD codes.
We calculated each ICD code's enrichment by dividing the observed proportion of honorees by the expected proportion of honorees and reported a value of log2 enrichment (LOE) and its 95% confidence intervals.
The 95% confidence interval of the LOE was estimated using the Poisson model method [@isbn:9780849394447].

Note that several ICD-9 codes have very small sample size (small number of patients associated with these codes).
Therefore, we should take that into account when interpreting the result of the statistical tests for these codes.

First, remove ICD with very small number of patients across site:

```{r}
small_icds <- diag_icd_9 %>% 
  group_by(icd) %>% 
  summarise(num_pats_icd = sum(num_patients_icd, na.rm = TRUE), .groups = 'drop') %>% 
  filter(num_pats_icd < 5) %>% 
  pull(icd)
diag_icd_9 <- diag_icd_9 %>% 
  filter(!icd %in% small_icds)
```

## Before admission

```{r warning=FALSE}
contingency_df <- diag_icd_9 %>% 
  # filter(time == 'Before admission') %>% 
  group_by(full_icd, time) %>% 
  summarise(across(contains('severe'), .fns = sum, na.rm = T), .groups = 'drop') %>% 
  mutate(Observed = num_patients_ever_severe_icd1, 
         num_non_severe = num_patients_never_severe_icd1 + num_patients_never_severe_icd0,
         num_severe = num_patients_ever_severe_icd1 + num_patients_ever_severe_icd0,
         Expected = num_patients_never_severe_icd1/num_non_severe*num_severe,
         over_sev = Observed - Expected)

nested_obs_exp <- contingency_df %>%
  select(full_icd, 
         time,
         num_patients_never_severe_icd0, 
         num_patients_never_severe_icd1, 
         num_patients_ever_severe_icd0, 
         num_patients_ever_severe_icd1) %>% 
  group_by(full_icd, time) %>% 
  nest() 

fish_obs_exp <- nested_obs_exp %>% 
  mutate(fish = map(data, my_fish)) %>% 
  dplyr::select(-data) %>% 
  unnest(cols = c(fish)) %>% 
  mutate(upper = ifelse(is.na(upper), Inf, upper),
         ci = paste0('(', round(log2(lower), 1), ', ', 
                     round(log2(upper), 1), ')'),
         lestimate = log2(estimate)) 
fish_obs_exp$`P value (Holm)` <-  p.adjust(fish_obs_exp$p_value, 'holm')
fish_obs_exp$`P value (FDR)` <- p.adjust(fish_obs_exp$p_value, 'BH')
```

While the Warning messages mentioned Chi-squared, the p-values were actually calculated using Fisher's Exact test (see more in `epitools::tab2by2.test()`).

Note: small number of observations for ICD-10 code G04, G03, G65:

```{r}
contingency_df %>% 
  filter(grepl('G04|G03|G65', full_icd)) %>% 
  datatable()
```


### Country enrichment table {#enrichment_tab}

The full table with all ICD codes and their corresponding enrichment can be browsed interactively below:
```{r}
library(DT)
fish_tab <- fish_obs_exp %>% 
  left_join(contingency_df, by = c('full_icd', 'time')) %>% 
  select(full_icd, time, Observed, Expected, over_sev,
         estimate, lestimate, ci, p_value,
         `P value (Holm)`, `P value (FDR)`) %>% 
  arrange(desc(over_sev)) %>% 
  rename('ICD' = 'full_icd',
         'Observed - Expected' = 'over_sev',
         'Enrichment' = 'estimate',
         'Log2(enrichment)' = 'lestimate',
         '95% Confidence interval' = 'ci', 
         'P value (raw)' = 'p_value')

fish_tab %>% 
  datatable(rownames = FALSE, filter = 'top') %>% 
  formatRound(c('Observed', 'Expected', 'Observed - Expected',
                'Enrichment', 'Log2(enrichment)'), 1) %>%
  formatSignif(c('P value (raw)', 'P value (Holm)', 'P value (FDR)'), 3) %>% 
  {.}

fish_tab %>% 
  write_csv('results/icd9_fish_tab.csv')

fish_obs_exp %>% 
  filter(`P value (FDR)` < 0.05) %>% 
  arrange(time) %>% 
  mutate(lestimate = round(lestimate, 2),
         drr = round((estimate - 1)*100, 0),
         lower_drr = (lower - 1)*100,
         upper_drr = (upper - 1)*100,
         ci_drr = paste0('(', round(lower_drr, 0), ', ', 
                     round(upper_drr, 0), ')'),
         p_fdr = format(`P value (FDR)`, digits = 2)) %>% 
  select(time, full_icd, lestimate, drr, ci_drr, p_fdr) %>% 
  write_csv('results/icd9_signi.csv')
```

A positive value of LOE indicates a higher proportion of *severe patients* with that ICD code compared to *non-severe patients*.
A LOE value of 1 represents a one-fold enrichment (i.e., observed number of *severe patients* is twice as much as expected).
We found an excess of *severe patients* with the following ICD codes:

- Other disorders of the brain (G93): 63 more *severe patients* than expected, LOE = [], 95% CI []
- Other and unspecified myopathies (G72): 38 more *severe patients* than expected, LOE = [], 95% CI []
- Myositis (M60): 6 more *severe patients* than expected, LOE = [], 95% CI []

### Compute enrichment from proportion comparisons

```{r fig.width = 7, fig.height = 3.5}
filtered_obs_exp <- contingency_df %>% 
  left_join(fish_obs_exp, by = c('full_icd', 'time')) %>% 
  mutate(
    distance_to_null = case_when(
      lower > 1 ~ lower - 1,
      TRUE ~ upper - 2
    ),
    presentation = case_when(
      lower > 1 & `P value (FDR)` < 0.05 ~ '#d46780', 
      upper < 1 & `P value (FDR)` < 0.05 ~ '#798234',
      TRUE ~ 'grey20'
    )) %>% 
  {.}
```

```{r warning=FALSE}
sorted_icds <- filtered_obs_exp %>% 
  filter(time == 'After admission') %>% 
  arrange(Expected) %>% 
  pull(full_icd)

plot_obs_exp <- filtered_obs_exp %>%
  mutate(lestimate = log2(estimate),
         llower = log2(lower), 
         lupper = log2(upper)) %>% 
  select(full_icd, time, lestimate, llower, lupper, presentation, over_sev, Observed, Expected) %>% 
  pivot_longer(- c(full_icd, time, presentation, over_sev), names_to = 'type') %>% 
  mutate(subtype = ifelse(type == 'Expected' | type == 'Observed', 
                          'Sqrt(number of honorees)', 
                          'Log2 enrichment, 95% CI')) %>% 
  pivot_wider(names_from = type) %>% 
  mutate(presentation = as.factor(presentation),
         full_icd = fct_relevel(full_icd, sorted_icds))

plot_obs_exp_right <- plot_obs_exp %>% filter(subtype == 'Sqrt(number of honorees)') 
plot_obs_exp_left <- plot_obs_exp %>% filter(subtype != 'Sqrt(number of honorees)') 
```


```{r fig.width = 12, fig.height = 6}
enrichment_plot_before <- plot_enrich(
  plot_obs_exp_left %>% filter(time == 'Before admission') %>% 
    mutate(presentation = droplevels(presentation)),
  plot_obs_exp_right %>% filter(time == 'Before admission') %>% 
    slice(match(sorted_icds, full_icd)) %>% 
    mutate(presentation = droplevels(presentation)),
  nudge = 1.5)

enrichment_plot_after <- plot_enrich(
  plot_obs_exp_left %>% filter(time == 'After admission') %>% 
    mutate(presentation = droplevels(presentation)),
  plot_obs_exp_right %>% filter(time == 'After admission') %>% slice(match(sorted_icds, full_icd)) %>% 
    mutate(presentation = droplevels(presentation)),
  nudge = 1.5)

enrichment_plot <- cowplot::plot_grid(
  enrichment_plot_before,
  enrichment_plot_after,
  ncol = 1, hjust = -0.05,
  labels = c('A. Before admission', 'B. After admission')
)
enrichment_plot
ggsave('figs/icd9_severe_after.png', enrichment_plot_after, width = 9.5, height = 3.5)
ggsave('figs/tiffs/efig_5.tiff', enrichment_plot_after, width = 9.5, height = 3.5, dpi = 300)
```

Figure caption: 
Each ICD code's log2 enrichment (LOE) and its 95% confidence interval (left), and the absolute difference between observed (triangle) and expected (circle) number of *severe patients* (right) before admission.
Positive value of LOE indicates a higher proportion of *severe patients* with that ICD code compared to *non-severe patients*.
Neurological ICD codes are ordered based on the number of *severe patients*.
Difference has been rounded.

```{r fig.width=9}
plot_enrich_both <- plot_obs_exp %>% 
  ggplot(aes(y = fct_relevel(full_icd, sorted_icds))) +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  ggstance::geom_pointrangeh(
    aes(x = lestimate,
        xmin = llower,
        xmax = lupper,
        color = time),
    position = position_dodge(width = 0.3), 
    stroke = 0.1, fatten = 3, size = 0.7  
  ) +
  labs(y = NULL, x = bquote(Log[2] ~ 'enrichment, 95% CI')) +
  theme(
    legend.position = 'bottom',
    legend.background = element_blank(),
    axis.title = element_text(size = 9),
    plot.margin = margin(5.5, 2, 5.5, 5.5, unit = 'pt')
  ) +
  scale_color_carto_d(palette = 4, direction = -1) +
  NULL
plot_enrich_both
ggsave('figs/icd9_severe.png', plot_enrich_both, width = 9, height = 5)
```
